---
title: "mahi"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mahi}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mahi)
```



```{r}
data(data_cont)
```

We ran simulations to validate MAHI and to compare it to methods recently introduced in the literature.

\subsection{Models}

\subsubsection{Continuous Outcome}

The dataset `data_cont` include $n = 100$ observations and $K = 500$ candidate mediators, according to the model

\begin{equation}
\begin{array}{ccl}
M_{ik} & = & \mu_k  + \alpha_k T_i + \epsilon_{ik}\\ 
Y_i & = & 20 + 50 T_i + \sum_k \beta_k M_{ik} +\epsilon_{i0}
\end{array}\label{grand_simed}
\end{equation}
where $1\leq i\leq n$ and $1\leq k\leq K$. The only exposure $T$ follows a Bernoulli distribution, $T \sim \mathcal{B}(0.3)$,  $\mu_k$ is drawn uniformly in the interval $[-2,2]$ for each variable $M_k$ and  $\epsilon_{k} \sim \mathcal{N}(0,1)$ for $k \in (0,\dots,500)$. Note that variables $M_k$ are causally unrelated one to the each other.

Table~\ref{grand_param} shows the values of $\alpha_k$ and $\beta_k$ for the first 30 variables $M_k$. The higher the absolute values of $\alpha_k,\beta_k$, the greater the indirect effect through $M_k$. As such, the first 10 mediators have strong indirect effects (and in principle are easier to select), the next 10 have mild indirect effects (less easy to detect) and the next 10 have weak indirect effects (hard to detect). All other 470 variables $M_k$ are not true mediators because either $\alpha_k=0$ or $\beta_k=0$.

\begin{table}[htp]
\begin{tabular}{|c|cccccccccc|}
\hline
$\boldsymbol{k}$ & \bf{1} & \bf{2} & \bf{3} & \bf{4} & \bf{5} & \bf{6} & \bf{7} & \bf{8} & \bf{9} & \bf{10} \\
\hline
$\boldsymbol{\alpha_{k}}$ & -95 & 90 & 95 & 95 & -100 & 95 & -95 & 85 & -95 & -100\\
$\boldsymbol{\beta_{k}}$ & 185 & -195 & 190 & 185 & -190 & 185 & 195 & -190 & 100 & 185\\
\hline
\hline
$\boldsymbol{k}$ & \bf{11} & \bf{12} & \bf{13} & \bf{14} & \bf{15} & \bf{16} & \bf{17} & \bf{18} & \bf{19} & \bf{20} \\
\hline
$\boldsymbol{\alpha_{k}}$ & -2.75 & 3.25 & -3.50 & 2.50 & -3.75 & 3.00 & -3.25 & 2.75 & 3.00 & 3.50\\
$\boldsymbol{\beta_{k}}$ & 3.25 & -2.50 & 3.75 & -3.00 & 3.50 & -2.75 & 3.75 & -3.25 & 3.25 & -2.75\\
\hline
\hline
$\boldsymbol{k}$ & \bf{21} & \bf{22} & \bf{23} & \bf{24} & \bf{25} & \bf{26} & \bf{27} & \bf{28} & \bf{29} & \bf{30} \\
\hline
$\boldsymbol{\alpha_{k}}$  & -0.875 & 0.625 & -0.375 & -0.25 & 0.50 & -0.875 & 0.125 & -1.125 & 0.375 &-0.25\\
$\boldsymbol{\beta_{k}}$  & 0.375 & -0.625 & 0.625 & -1.125 & 0.75 & 0.25 & -0.50 & 0.375 & -0.625 & 0.75\\
\hline
\hline
$\boldsymbol{k}$& \bf{31} & \bf{32} & \bf{33} & \bf{34} & \bf{35} & \bf{36} & \bf{37} & \bf{38} & \bf{39} & \bf{40} \\
\hline
$\boldsymbol{\alpha_{k}}$ & 25 & 25 & 25 & 25 & 25 & 25 & 25 & 25 & 25 & 25\\
$\boldsymbol{\beta_{k}}$ & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
\hline
\hline
$\boldsymbol{k}$ & \bf{41} & \bf{42} & \bf{43} & \bf{44} & \bf{45} & \bf{46} & \bf{47} & \bf{48} & \bf{49} & \bf{50} \\
\hline
$\boldsymbol{\alpha_{k}}$  & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
$\boldsymbol{\beta_{k}}$  & 45 & 45 & 45 & 45 & 45 & 45 & 45 & 45 & 45 & 45\\
\hline
\end{tabular}\caption{
Values of $\alpha_k$ and $\beta_k$ for $k=1,\ldots,50$. For $k=51,\ldots,500$, $\alpha_k=\beta_k=0$.} \label{grand_param}
\end{table}


# First Step

Before using the function mahi we have to choose a value for lambda using lambdaChoice function. We have to set a max for lambda with lambdamax. The returned value is less than lambdamax and allows you to select a mediator number between selectedMin and selectedMax. The value of selectedMax must be less than or equal to the number of individuals n. If the value of lambdamax is not high enough, a new lambdamax is test by adding 50 to previous value. 

```{r}
lambda <- lambdaChoice(data_cont,name.exposure="Exposure",name.outcome="Outcome",name.mediators=paste0("M", 1:500),lambdamax=30600,selectedMin=50,selectedMax=100)
```

Now you can use the mahi function for the selection
```{r,include=FALSE}
  mahifit <- mahi(data=data_cont,name.exposure="Exposure",name.outcome="Outcome",name.mediators=paste0("M", 1:500),
                   lambda=lambda,Nboot=30,L0=.1,eta=2,epsilon=.001,
                   Kmax=30,bin=FALSE,
                   dostep2=TRUE,p.adjust.method="hochberg",pvalseuil=0.05)
```

```{r,eval=FALSE}
  mahifit <- mahi(data=data_cont,name.exposure="Exposure",name.outcome="Outcome",name.mediators=paste0("M", 1:500),
                   lambda=lambda,Nboot=30,L0=.1,eta=2,epsilon=.001,
                   Kmax=30,bin=FALSE,
                   dostep2=TRUE,p.adjust.method="hochberg",pvalseuil=0.05)
```


```{r}
summary(mahifit)
```

The function mahi returns a list containing several results such as the mediators selected in the first step. 


To see the mediators selected after the step 1, use :
```{r}
  mahifit$Kmaxrank
```
Remember that the first 30 candidates are the real mediators. The first 10 (1 to 10) have a strong mediated effect, the next 10 (11 to 20) have a medium mediated effect, and the other 10 (21 to 30) have a very weak effect.


To see the mediators count for each mediator after the stability selection (step 1), use :
```{r}
  mahifit$bootcount
```

To see the ranking after the stability selection (step 1), use :
```{r}
mahifit$ranking
```


To see the results of the multiple analysis with mediators selected at the step 1, use :
```{r}
  summary(mahifit$multimed)
```

You can see the p-value in the previous table or use :
```{r}
  mahifit$pvals
```


The corrected p-value for the multiple test obtained with method given can be disp with :
```{r}
  mahifit$pvalscorr
```
It is then possible to make another correction for a multiple test with the mediators selected in step 1.

To see the mediators selected after step 2 you can use these lines of codes.
```{r}
mahifit$step2
```
